Canonical rings of log spin stacky curves

### 2015-07-08 curves-write-up/spin_cring.tex
added generation 4

Theorem ~\ref{thm:main} is proven separately in the cases that the genus $g = 0, g = 1,$ and $g \geq 2$ in Theorems ~\ref{thm:g-0-main}, ~\ref{thm:g-1-main}, and ~\ref{thm:g-high-main}, respectively.  In each of these proofs, we follow a similar inductive process utilizing the lemmas of Section ~\ref{sec:induction}; however, in the first two we have to explicitly construct specific base cases and present a finite list of exceptional cases, while in the genus $g \geq 2$ case we deduce base cases from more general arguments.
\begin{rem}
\label{rem:explicit-generators}
The main theorem obtains bounds on the degrees of the generators
and relations of log spin canonical rings. However, the proof
of the genus one and genus zero cases also yield explicit systems
of generators and an explicit initial ideal of relations, as described in Remarks ~\ref{rem:genus-1-explicit-bound} and ~\ref{rem:genus-0-explicit-bound}.
Furthermore, our proof of the genus $g \geq 2$ case also provides
an inductive procedure for explicitly determining the generators
and initial ideal of relations of log spin canonical rings. Many
explicit systems of generators and relations for curves of genus $2
\leq g \leq 15$ are detailed in interesting examples by Neves
\cite[Section III.4]{neves:halfcan}. 
\end{rem}
\begin{rem}
The explicit construction described in Remark ~\ref{rem:explicit-generators} also reveals that the bounds given in Theorem ~\ref{thm:main} are tight.
Specifically, in almost all cases, the log spin canonical ring requires a generator in degree $e = \max(5, e_1, \ldots, e_r)$ and a relation in degree at least $2e-4$. 
Furthermore, there are many infinite families of cases which require a generator in degree $e = \max(5, e_1, \ldots, e_r)$ and a relation in degree exactly $2e$. For further detail, see Remarks ~\ref{rem:genus-0-explicit-bound}, ~\ref{rem:genus-1-explicit-bound}, and ~\ref{rem:genus-2-explicit-bound} in the cases that the genus is $0,1,$ or $\geq 2$ respectively.
\end{rem}

@@

\begin{rem}
\label{rem:gen-at-most-three}
If $M(\Gamma)$ has some odd weight modular form, then it has an odd weight modular form in weight $3$. When $g \geq 2$, we see that this is true because $\dim_k H^0(\sx, 3\halfcan) > 0$ by Riemann--Roch and the fact that $\deg \lfloor 3\halfcan \rfloor > 2g - 1$. When the genus is zero or one, we see that there is a generator in weight $1$ or weight $3$ in the base cases given in Table ~\ref{table:g-1-base} and Table ~\ref{table:g-0-base-cases} so there is an odd weight modular form in weight $3$ in general. A consequence of this observation is that the bound on the degree of generators and relations when $M(\Gamma)$ has some odd weight modular form, as given in Corollary ~\ref{cor:main-mod-forms}, is closely related to the degree of the minimal odd weight modular form.
\end{rem}

\begin{example}
\label{eg:congruence-bounds}
Suppose $\Gamma \subset SL_2(\BZ)$ is a congruence subgroup. Then, since 
the action  of $SL_2(\BZ)$ has points with stabilizer order 2 and 3, the 
action of $\Gamma$ can only have points with stabilizer order 2 and 3. 
Furthermore, if $\Gamma$ has some odd weight modular form, then it must 
have no points with stabilizer order 2 (see Remark ~\ref{rem:odd-denom}). 
Therefore, by Corollary ~\ref{cor:main-mod-forms}, $M(\Gamma)$ is 
generated in degree at most $5$ with relations in degree at most $12$. 
Furthermore, if the coarse space associated $\Gamma$ does not have, 
$(0;3,3,3,3;0),$ there are relations in degree at most $10$, as follows 
from the list of exceptional cases in Tables ~\ref{table:genus-0-exceptional-cases} and ~\ref{table:genus-0-exceptional}.
\end{example}

@@

\begin{rem}
\label{rem:genus-2-explicit-bound}
Using the terminology from the proof of Theorem ~\ref{thm:g-high-main}, if $e = \max(e_1,\ldots, e_r) \geq 5$, we see that $R_L,$ has a generator in precisely $e$. This follows from the inductive application of Lemmas ~\ref{lem:sat-1} and \ref{lem:sat-2}, which, at the last step, add a minimal generator of $R_L$ over $R_L'$ in degree $e$. Similarly, this inductive application shows that our given presentation has a relation in degree $2e-4$.
\end{rem}

@@

\begin{rem}
\label{rem:genus-1-explicit-bound}
In addition to the bound on the degree of the generators and relations, as detailed in Theorem ~\ref{thm:g-1-main}, proof of Theorem ~\ref{thm:g-1-main} yields an explicit procedure for computing those minimal generators and relations. That, is one can start with the generators and relations found in the base cases, and inductively add in generators and relations as one adds and then increments stacky points. Since the proof of Theorem ~\ref{thm:g-1-main} is given by inductively applying Lemmas ~\ref{lem:sat-1} and ~\ref{lem:sat-2}, when $e = \max(5,e_1,\ldots, e_r) \geq 5$ we obtain that $R_L$ has a generator in degree $e$ and a relation in degree at least $2e - 4$. Further, by examining the statements of Lemma ~\ref{lem:sat-1} and \ref{lem:sat-2} in the case that there $1 \leq i < j \leq r$ with $e_i = e_j = e$, then, at least in our construction, there is a relation in degree $2e$.
\end{rem}

@@

\begin{proof}
If $\sx$ has no stacky points, then we can assume that $\halfcan \sim
n \cdot \infty$ with $n \in \BZ_{\geq -1}$. This is a classical
case which is done by Voight and Zureick-Brown \cite[Section 4.2]
{vzb:stacky}. When $n = -1$, then $R_\halfcan = \Bk$.  When $n = 0$,
then $R_\halfcan = \Bk[x]$.  When $n > 0$, inductively applying Lemma
\ref{lem:sat-1} tells us that $R_\halfcan$ is generated in degree 1
with relations generated in degree 2.

@@

\begin{rem}
\label{rem:genus-0-explicit-bound}
Just like the case of genus 1, as described in Remark ~\ref{rem:genus-1-explicit-bound}, the proof of Theorem ~\ref{thm:g-0-main} gives an explicit construction of the generators and relations for the log spin canonical ring $R_L$. Furthermore, similarly to the description given in Remarks in ~\ref{rem:genus-2-explicit-bound} and Remark ~\ref{rem:genus-0-explicit-bound}, when $e = \max(e_1, \ldots, e_r) \geq 5,$ there is a generator in degree $e$ and a relation in degree at most $2e$. This can be seen from the inductive application of Lemmas ~\ref{lem:sat-1}, \ref{lem:sat-2}, and \ref{lem:sat-3}. Additionally, as in the case of genus 1, if there are $1 \leq i < j \leq r$ with $e = e_i = e_j$ then there is a relation in degree $2e$. 
\end{rem}

>>>>

Theorem ~\ref{thm:main} is proven separately in the cases that the genus $g = 0, g = 1,$ and $g \geq 2$ in Theorems ~\ref{thm:g-0-main}, ~\ref{thm:g-1-main}, and ~\ref{thm:g-high-main}, respectively.  In each of these proofs, we follow a similar inductive process utilizing the lemmas of Section ~\ref{sec:induction}; however, in the first two we have to explicitly construct specific base cases and present a finite list of exceptional cases, while in the genus $g \geq 2$ case we deduce base cases from more general arguments.
\begin{rem}
\label{rem:explicit-generators}
The main theorem obtains bounds on the degrees of the generators
and relations of log spin canonical rings. However, the proof
of the genus one and genus zero cases also yield explicit systems
of generators and an explicit initial ideal of relations, as described in Remarks ~\ref{rem:genus-1-explicit-bound} and ~\ref{rem:genus-0-explicit-bound}.
Furthermore, our proof of the genus $g \geq 2$ case also provides
an inductive procedure for explicitly determining the generators
and initial ideal of relations of log spin canonical rings from a presentation of the canonical ring of the coarse space. Many
explicit systems of generators and relations for curves of genus $2
\leq g \leq 15$ are detailed in interesting examples by Neves
\cite[Section III.4]{neves:halfcan}. 
\end{rem}
\begin{rem}
The explicit construction described in Remark ~\ref{rem:explicit-generators} also reveals that the bounds given in Theorem ~\ref{thm:main} are tight.
Specifically, in almost all cases, the log spin canonical ring requires a generator in degree $e = \max(5, e_1, \ldots, e_r)$ and a relation in degree at least $2e-4$. 
Furthermore, there are many infinite families of cases which require a generator in degree $e = \max(5, e_1, \ldots, e_r)$ and a relation in degree exactly $2e$. For further detail, see Remarks ~\ref{rem:genus-0-explicit-bound}, ~\ref{rem:genus-1-explicit-bound}, and ~\ref{rem:genus-2-explicit-bound} in the cases that the genus is $0,1,$ or $\geq 2$ respectively.
\end{rem}

@@

\begin{rem}
\label{rem:gen-at-most-three}
If $M(\Gamma)$ has some odd weight modular form, then it has an odd weight modular form in weight $3$. When $g \geq 2$, we see that this is true because $\dim_k H^0(\sx, 3\halfcan) > 0$ by Riemann--Roch and the fact that $\deg \lfloor 3\halfcan \rfloor > 2g - 1$. When the genus is zero or one, we see that there is a generator in weight $1$ or weight $3$ in the base cases given in Table ~\ref{table:g-1-base} and Table ~\ref{table:g-0-base-cases}. Hence, there is an odd weight modular form in weight $3$ in general. A consequence of this observation is that the bound on the degree of generators and relations when $M(\Gamma)$ has some odd weight modular form, as given in Corollary ~\ref{cor:main-mod-forms}, is closely related to the degree of the minimal odd weight modular form.
\end{rem}

\begin{example}
\label{eg:congruence-bounds}
Suppose $\Gamma \subset SL_2(\BZ)$ is any congruence subgroup. Then, since
the action  of $SL_2(\BZ)$ on $\BH$ only has points with stabilizer order 1, 2 and 3, and has a cusp, the 
action of $\Gamma$ can only have points with stabilizer order 1, 2 and 3, and has at least one cusp.
Furthermore, if $\Gamma$ has some odd weight modular form, then it must 
have no points with stabilizer order 2 (see Remark ~\ref{rem:odd-denom}). 
Therefore, by Corollary ~\ref{cor:main-mod-forms}, $M(\Gamma)$ is 
generated in degree at most $5$ with relations in degree at most $10$. In fact, it is not difficult to reduce generation in weight at most 5 to generation in weight at most 4, as noted in Remark ~\ref{rem:cusp-generation-4}. Note that the exceptional cases in Tables ~\ref{table:g-1-exceptional} and ~\ref{table:g-0-exceptional} do not occur because there is a cusp, so $\delta > 0$.
\end{example}

@@

\begin{rem}
\label{rem:genus-2-explicit-bound}
For this remark, we retain the terminology from the proof of Theorem ~\ref{thm:g-high-main}.
Suppose $e :=  \max(e_1,\ldots, e_r).$ Then, $R_L$ has a generator in degree $e$ when $e \geq 5$ and a relation in degree at least $2e-4$ when $e \geq 7$. Since the proof of Theorem ~\ref{thm:g-1-main} is given by inductively 
applying Lemmas ~\ref{lem:sat-1} and ~\ref{lem:sat-2}, we obtain that $R_L$ is minimally generated over $R_{L'}$ by an element in degree $e,$ assuming $e \geq 5$. Furthermore, if $e \geq 7$ and $e_i = e,$ 
then there must be a relation with leading term $y_{i,e_i} \cdot y_{i,e_i-4}.$ Hence, there is a relation in degree at least $2e - 4$. 
Further, by examining the statements of Lemma ~\ref{lem:sat-1} and \ref{lem:sat-2} in the case that there are $1 \leq i < j \leq r$ with $e_i = e_j = e$, then, there is necessarily a relation with leading term $y_{i, e_i} \cdot y_{j, e_j}$ in degree $2e$.

This analysis also applies to the cases that $g = 0$ and $g = 1$.
\end{rem}

@@

\begin{rem}
\label{rem:genus-1-explicit-bound}
In addition to the bound on the degree of the generators and relations, as detailed in Theorem ~\ref{thm:g-1-main}, the proof of Theorem ~\ref{thm:g-1-main} yields an explicit procedure for computing those minimal generators and relations. That, is one can start with the generators and relations found in the base cases and inductively add in generators and relations as one adds and then increments stacky points. As described in Remark ~\ref{rem:genus-2-explicit-bound}, when $e: = \max(e_1,\ldots, e_r) \geq 7$, there is necessarily a generator in degree $e$ and a relation in degree $2e$.
\end{rem}

@@

\begin{proof}
If $\sx$ has no stacky points, then we can assume that $\halfcan \sim
n \cdot \infty$ with $n \in \BZ_{\geq -1}$. This is a classical
case done by Voight and Zureick-Brown \cite[Section 4.2]
{vzb:stacky}. When $n = -1$, then $R_\halfcan = \Bk$.  When $n = 0$,
then $R_\halfcan = \Bk[x]$.  When $n > 0$, inductively applying Lemma
\ref{lem:sat-1} tells us that $R_\halfcan$ is generated in degree 1
with relations generated in degree 2.

@@

\begin{rem}
\label{rem:genus-0-explicit-bound}
Similarly to the case of genus 1, as described in Remark ~\ref{rem:genus-1-explicit-bound}, the proof of Theorem ~\ref{thm:g-0-main} gives an explicit construction of the generators and relations for the log spin canonical ring $R_L$. Furthermore, as described in Remark ~\ref{rem:genus-2-explicit-bound}, when $e := \max(e_1, \ldots, e_r)$ is at least $7$, there is a generator in degree $e$ and a relation in degree at least $2e-4$. This can be seen from the inductive application of Lemmas ~\ref{lem:sat-1}, \ref{lem:sat-2}, and \ref{lem:sat-3}.
\end{rem}

\begin{rem}
\label{rem:cusp-generation-4}
Here, we describe how to obtain a slightly better bound for our application to modular forms from example ~\ref{eg:congruence-bounds}. A careful scrutiny of Theorem ~\ref{thm:g-0-main} reveals that, if $\Delta > 0$ and $\sx$ has signature $(0;3, \ldots, 3; \delta)$, then $R_L$ is generated in degree at most 4. This is $\delta > 0$, so $L_X$ is effective. Addtionally, $R_{L_X}$ is generated in degree 1, and inductively applying Lemma ~\ref{lem:sat-1} only adds generators in degree 3 and 4. Therefore, $R_L$ is generated in degree at most 4. Note that a similar analysis of the proof of Theorem ~\ref{thm:g-1-main} yields that when $g = 1$, congruence subgroups are generated in degree at most 4.
\end{rem}
